---
title: "Attribution Workshop Bug Analyses"
author: "Emma Jones"
output: html_notebook
---
  
  This notebook walks users through the process of organizing data and running the CDF, GAM, and TITAN models on VDEQ bug data (and data from other entities) for the VDEQ species attribution workshop.

Data processing
```{r,echo=F,message=F}
library(tidyverse)
library(plotly)
library(reshape)
library(Hmisc)
library(mgcv)
#library(TITAN2)
```

Read in taxa data. Split dataset into environmental and taxa information. This dataset was created by merging all EDAS bug data with probmetrics dataset 2001-2014.
```{r, echo=FALSE}
bugData <- read_csv('data/processedData/allSpringBugs.csv')
bugData <- as.data.frame(bugData)
y <- filter(bugData,UID %in% bugData[duplicated(bugData$UID),]$UID)
# get rid of Fairfax QA data bc not unique UID
bugData <- filter(bugData,!(UID %in% y$UID));rm(y)
# Then change UID back to original format so it can match to the envData (which has no way of getting a BenSampID)
#bugData$UID <- gsub("\\/.*","",bugData$UID)
#y <- filter(bugData,UID %in% bugData[duplicated(bugData$UID),]$UID)%>%select(UID,StationID,CollDate,BenSampID)
# drop weird reps by BenSampID, these were chosen by first date (where available) or lower bensampid
#bugData <- filter(bugData,!(BenSampID %in% 
#                     c('POH1727','XMJ9228','cks7171','stm9745','bmt7502','DCK2262',
#                       'ram7608','SCM4727','XRA7458','XRD7463','YST7460','YST7461',
#                       'YST7462','rap7607','RNF3593','ROA8012','TKR9147','wrn4827',
#                       'XOK7455','XOK7456','XOK7457','MHN6030','PBXATTRO1b',
#                       'KOX8266','CPN8128','RMNRAM6287','BLD1827')))
bugTaxa <- select(bugData,UID,Acentrella:Zygoptera)
bugTaxa <- as.data.frame(bugTaxa)
bugTaxa[2:491] <- apply(bugTaxa[2:491], 2, function(x) as.numeric(as.character(x)))
bugTaxa[is.na(bugTaxa)]=0


envData <- read_csv('data/processedData/VA_FF_WVA_MD_Other_envData.csv') %>%
  select(UID,DataSource,StationID,Year,LongitudeDD,LatitudeDD,totalArea_sqMile,
         Basin,SubBasin,EcoRegion,BioRegion,DO,pH,SpCond,TN,TP,TotHab,
         TotHab_wMD,Cl,Sf, ## PARAMETERS ADDED FOR ROUND TWO OF ANALYSES
         LRBS,MetalCCU,wshdImpPCT)
envData[envData=="NA"]=NA
envData[12:23] <- apply(envData[12:23], 2, function(x) as.numeric(as.character(x)))
envData <- as.data.frame(envData)



# change all / in bug taxa names to _ so doesnt mess up file directory structure with paste()
names(bugTaxa) <- gsub(x = names(bugTaxa), pattern = "/", replacement = "_")

```

```{r envDataQA, echo=FALSE}
# Get rid of odd measures
weirdDO <- filter(envData, DO > 14) 
weirdpH <- filter(envData, pH > 14 & pH < 0) 
weirdSpCond <- filter(envData, SpCond > 5000)
weirdTN <- filter(envData, TN > 5)
weirdTP <- filter(envData, TP > 1)
weirdTotHab <- filter(envData, TotHab < 0 | TotHab > 200)
weirdTotHab_wMD <- filter(envData, TotHab_wMD < 0 | TotHab_wMD > 200)
weirdLRBS <- filter(envData, LRBS < -4 | LRBS > 2)
weirdwshdImpPCT <- filter(envData, wshdImpPCT < 0 | wshdImpPCT > 100)
rm(weirdpH);rm(weirdSpCond);rm(weirdTN);rm(weirdTP);rm(weirdTotHab);rm(weirdTotHab_wMD);rm(weirdLRBS);rm(weirdwshdImpPCT)

# if these are worth dropping, then do next step 
sitesToDrop <- c(weirdDO$UID)#,weirdpH$UID,#weirdSpCond$UID,weirdTN$UID,weirdTP$UID,weirdTotHab$UID,weirdLRBS$UID,weirdwshdImpPCT$UID)

envData <- filter(envData, !(UID %in% sitesToDrop))


```



### CDF
Function to output calculate cumlative percentiles for any input dataset.
```{r, echo= F}
# Testing parameters
#dataset <- fishTaxa
#parameter <- select(envData,UID,DO)
#parameter <- 'DO'
#taxaAndEnvdataset <-  merge(fishTaxa,select(envData,UID,DO),by='UID')
  
CDF <- function(taxaAndEnvdataset,parameter){
  #dataset <- merge(dataset,parameter,by='UID')
  dataset <- taxaAndEnvdataset[order(taxaAndEnvdataset[, parameter]), ]
  value <- dataset[, parameter]
  taxa.names <- sort(names(select(dataset,-UID)))
  for(i in 1:length(taxa.names)){
    csum <- cumsum(dataset[,taxa.names[i]]) / sum(dataset[,taxa.names[i]])
  dataset <- cbind(dataset,csum)
  names(dataset)[length(taxa.names)+1+i] <- paste('csum_',taxa.names[i],sep="")}
  dataset <- mutate(dataset,parameter=parameter, value=value) %>%
    select(UID,parameter,value,everything())
  return(dataset)
}

# How to run function
#cdfdata <- CDF(dataset,select(envData,UID,DO))
#cdfdata <- CDF(merge(fishTaxa,select(envData,UID,DO),by='UID'),'DO')

# Testing Parameters
#CDFdataset <- cdfdata
#taxaName <- "Acantharchus  pomotis"
#percentile <- 0.75
#options(warn = -1)

CDFplot <- function(CDFdataset,taxaName){
  # Double spaces are a pain
  keep <- c('parameter','value', paste('csum_',taxaName,sep=""))
  ggData <- CDFdataset[,match(keep,names(CDFdataset))] %>%
    dplyr::rename(pct=!!names(.[3]))
  plot_ly(ggData) %>%
    add_trace(x = ~value, y = ~pct, mode = 'scatter', name=taxaName,
                  hoverinfo="text",text=~paste(format(pct,digits=2)," , ",format(value,digits=2))) %>%
    layout(showlegend = FALSE,
         title=taxaName,
         xaxis = list(title = "Proportion of Total"),
         yaxis = list(title = ggData$parameter[1]))
   #ggplot(ggData,aes(value,pct)) + geom_line() +
   # theme(plot.title = element_text(size=20, face="bold", margin = margin(10, 0, 10, 0))) + 
    #labs(x=ggData$parameter[1],y="Proportion of Total",title=taxaName) 
 
} 

# How to run function
#CDFplot(cdfdata,names(cdfdata)[15])

```


Nonparametric Regression- Generalized Additive Model. This function is adapted from Lei Zheng (lei.zheng@tetratech.com) scripts for plots with the GAM and relative abundace.
```{r, echo= F}
options(scipen=10)


tolerance <- function(spdata, envdata,  sp.siteid="Sample.ID", species="GENUS", 
                      sp.abndid="RA", env.siteid="Sample.ID", xvar="cond", 
                      cutoff = 20, cutoff2 = 10, region = "all", lim ="GAM",
                      coord = NULL, mtype = 3, dense.N = 201, cast = TRUE,  
                      plot.pdf = F, add.map=F,statename = NULL,  add.lab = F, 
                      add.abund = T, ylab1 = "Capture Probability",  
                      ylab2 = "Relative Abundance", 
                      main = "Capture Probability of Macroinvertebrate Taxon Along Conductivity Gradient", 
                      mar = c(5,4,3, 4), 
                      xlabs=expression(paste("Conductivity ( ", mu, "S/cm)")), 
                      log.x = TRUE, base = 10, plus = F, rounder = 2, taus = c(50,95), 
                      nbin = 60, emmaLOG = FALSE) {
  
  print(xvar)
  dfenv <- envdata[!is.na(envdata[,xvar]), c(env.siteid, xvar, coord)] 
  if(cast) { 
    df.sp <- data.frame(SITEID = spdata[, sp.siteid], species = spdata[, species],
                        abund = spdata[, sp.abndid]) 
    df.sp <- subset(df.sp, !is.na(species) & !is.na(abund))
    df.tmp <- merge(df.sp, dfenv, by.x = "SITEID", by.y = env.siteid )
    ss <- reshape::cast(df.tmp, SITEID ~ species, sum, value = "abund") 
    names(ss)[1] <- sp.siteid
  } else ss <- spdata
  #    print(names(ss))
  df1 <- merge(ss, dfenv, by.x = sp.siteid, by.y = env.siteid)   # merged crosstab abundance file
  nc <- ncol(df1)
  taxa.count <- apply(df1[,(2:ncol(ss))] > 0, 2, sum)
  #    print((taxa.count))
  tnames <- names(taxa.count)[taxa.count >= cutoff2] 
  ntaxa <- length(tnames)                           # length of taxa list for taxonomic level i
  
  df2 <- df1 <- df1[order(df1[,xvar]),]
  
  df2[2:(nc-ncol(dfenv) + 1)] <- apply(df2[2:(nc-ncol(dfenv) + 1 )] > 0, 2, as.numeric)                 # p/a file
  
  stress <- df1[, xvar]
  stress.u <- sort(unique(stress))
  
  xrange <- range(stress)
  xnew <- seq(from = xrange[1], to = xrange[2], length = 100)
  
  ###### calculate weight for cdf 
  
  cutp <- seq(from = min(stress), to = max(stress), length = nbin)
  cutm <- 0.5*(cutp[-1] + cutp[-nbin])
  
  df2$cutf <- cut(df2[,xvar], cutp, include.lowest = T)
  
  wt <- 1/table(df2$cutf)
  wt[is.infinite(wt)] <- NA
  wt <- wt/sum(wt, na.rm = T)
  dftemp <- data.frame(cutf = names(wt), wt = as.vector(wt))
  df2 <- merge(df2, dftemp, by = "cutf")[-1]
  df2 <- df2[order(df2[,xvar]),]  
  df1 <- df1[order(df1[,xvar]),] 
  
  #### end 
  ##### function to compute area under the curve
  auc <- function(xrange, mod, dense.N) {
    x <- seq(min(xrange), max(xrange), length = dense.N - 1)
    s.area <-rep(NA, dense.N - 1)
    y <- predict(mod, newdata = data.frame(dose = x), type = "response")
    for (index in 1:dense.N-1) {
      s.area[index] <- (y[index] + y[index + 1])/2*(x[index+1]-x[index])
    }
    tsum <- sum(s.area, na.rm=T)         # all area
    jj=1
    csum = sum(s.area[1:jj])              # cum area
    xc95 <- yc95 <- rep(NA, length(taus))
    for(ii in 1:length(taus)) {
      while(csum < taus[ii]/100*tsum) {
        jj = jj + 1
        csum <- sum(s.area[1:jj], na.rm=T)
      }
      if(jj == 1) {
        xc95[ii] <- x[jj]
        yc95[ii] <- y[jj]
      } else {
        xc95[ii] <- (x[jj]+ x[jj-1])/2  
        yc95[ii] <- (y[jj]+ y[jj-1])/2
      }
    }
    return(c(xc95))
    #      print(xc95)
  } 
  # plot pdf option
  #if(plot.pdf) {
  #  pdf(file = paste("data/", region, ".taxon.gam.pdf",sep=""),
  #      width = 9, height = 6.5, pointsize = 12)
  #  par(mfrow = c(2, 3), pty = "m", mar =mar, oma=c(1.5, 0.5, 2,0.5) )
  #  totpage <- ceiling((length(tnames)+1)/6)
  #  
  #  if(add.map) {
  #    simpleCap <- function(x) {
  #      s <- strsplit(x, " ")[[1]]
  #      paste(toupper(substring(s, 1,1)), substring(s, 2),
  #            sep="", collapse=" ")
  #    }
  #    med = 1
  #    while(is.na(df1[med, coord[1]])) med = 1 + med
  #    
  #    if(is.null(statename)) {
  #      statename <<- map.where("state", df1[med, coord[1]], df1[med, coord[2]])
  #      statename <- simpleCap(statename)
  #    } 
  #    if(add.lab) {
  #      map.text("state", region = statename, mar = mar)
  #    } else {
  #      map("state", regions = statename, mar = mar)
  #    }
  #    
  #    text(df1[,coord[1]],df1[,coord[2]],".")
  #    #       mtext(paste("Ecoregion",region), side =3,line = 0.5, cex=1.5)  
  #  }
  #}
  
  varnames <- c("N", "min_ob", paste(taus, "th_ob",sep="_"), "max_ob", "Opt_WA",
                "Tol_WA", paste("CDF", taus,"th","Abund", sep ="_"),
                paste("CDF", taus,"th","PA", sep ="_"),paste("CDF_wt", taus,"th",sep ="_"), 
                paste("LRM", taus,"th", sep ="_"),paste("QLRM", taus,"th", sep ="_"),
                "Opt_qlrm", "Tol_qlrm", paste("GAM",taus, "th", sep="_"), "ROC")
  
  optsave <- matrix(NA, ncol = length(taus)*7 + 8, nrow  = ntaxa)
  colnames(optsave) <- varnames
  rownames(optsave) <- tnames
  
  for (i in 1:ntaxa) {
    isel <- match(tnames[i], names(df1))       # selected taxa i
    #    print(df1[,isel])
    print(tnames[i])
    ### (0) observed range
    samplen <- length(df2[df2[,isel]>0,xvar])          
    limits <- quantile(df2[df2[,isel]>0,xvar], prob = c(0, taus/100, 1))       
    
    ## (1)  Weighted averaging 
    WA <- sum(df1[,isel]*df1[,xvar])/sum(df1[,isel]) 
    tol <- sqrt(sum(df1[,isel] * (df1[,xvar]- WA)^2) /sum(df1[,isel])) 
    #    WTA <- sum(df1[,isel]*df1[,xvar]*df2$wt)/sum(df1[,isel]*df2$wt)     # sample weighted weighted average         
    
    ### (2)  cdf or cumsum no weighting  both abundance based and p/a based
    csum1 <-cumsum(df1[,isel])/sum(df1[,isel])          # cumlative vectors devided total abundance 
    #      print(sum(df1[,isel]))
    ic1 <- rep(1, length(taus))
    for(ii in 1:length(taus)) {
      #  print(csum1)
      if(is.finite(csum1[ic1[ii]])) {
        while(csum1[ic1[ii]] < taus[ii]/100) ic1[ii] <- ic1[ii] + 1
      } else csum1 <- NA
    }
    #### (3) cdf p/a        
    csum2 <-cumsum(df2[,isel])/sum(df2[,isel])    # cumlative vectors devided total abundance 
    ic2 <- rep(1, length(taus))
    for(ii in 1:length(taus)) {
      while(csum1[ic2[ii]] < taus[ii]/100) ic2[ii]<- ic2[ii] + 1
    }
    
    ##### (4) cdf weighted 
    pres <- df2[, isel] > 0
    sel <- df2[pres,]
    eout2 <- rep(NA, length(taus))
    for(ii in 1:length(taus)) {
      eout2[ii] <- Hmisc::wtd.quantile(sel[,xvar], sel$wt, normwt = TRUE, prob = taus[ii]/100)  
    }
    
    ######### (5)(6)(7) logistic regression model          
    resp <<- df2[, isel] 
    dose <<- df2[, xvar]
    lrm1 <- glm(resp ~ dose, family="binomial")
    lrm2 <- glm(resp ~ dose + I(dose^2), family="binomial")
    #### u = -b1/(2b2)  t = 1/sqrt(-2b2) for lrm2  
    lrm2.opt <- (-lrm2$coef[2]/(2*lrm2$coef[3])  )
    lrm2.tol <- 1/sqrt(-2*as.vector(lrm2$coef[3]))
    s=mgcv:::s
    
    lrm3 <- try(mgcv::gam(resp ~ s(dose,k = 3), family= "binomial"),TRUE)
    
    if(isTRUE(class(lrm3)=="try-error")) { next } else{  ############################ ERROR HANDLING INSERTED BY EVJ 02062018
    
    if(mtype == 1) model <- lrm1
    if(mtype == 2) model <- lrm2
    if(mtype == 3) model <- lrm3

    # Compute mean predicted probability of occurrence
    predlk <- predict(model, type = "link")
    predout <- exp(predlk)/(1 + exp(predlk))
    x <- predout[pres]
    y <- predout[!pres]
    rocmat <- matrix(NA, nrow = length(x), ncol = length(y))
    for (j in 1:length(x)) {
      rocmat[j,] <- as.numeric(x[j] > y)
    }
    # Summarize all comparisons to compute area under ROC
    roc <- sum(rocmat)/(length(x)*length(y))
    #        wilcox <- wilcox.test(x, y, conf.int = FALSE)
    predresp <- predict(model, newdata = data.frame(dose=xnew), type="link", se.fit=T)
    
    # Compute upper and lower 90% confidence limits
    up.bound.link <- predresp$fit + qnorm(0.95)*predresp$se.fit
    low.bound.link <- predresp$fit + qnorm(0.05)*predresp$se.fit
    mean.resp.link <- predresp$fit
    
    # Convert from logit transformed values to probability.
    up.bound <- exp(up.bound.link)/(1+exp(up.bound.link))
    low.bound <- exp(low.bound.link)/(1+exp(low.bound.link))
    mean.resp <- exp(mean.resp.link)/(1+exp(mean.resp.link))
    
    #### 5,6,7 full range
    ######## find modeled xc95 full data range
    lrm1.95f <- auc(xrange = xrange, mod = lrm1, dense.N = dense.N)
    lrm2.95f <- auc(xrange = xrange, mod = lrm2, dense.N = dense.N)
    lrm3.95f <- auc(xrange = xrange, mod = lrm3, dense.N = dense.N) 
    #        print(lrm3.95f)
    ###### Save results 
    
    optsave[i,] <- c(samplen, round(c(limits, WA, tol, df1[ic1,xvar],
                                      df1[ic2,xvar], eout2, lrm1.95f,lrm2.95f, lrm2.opt, lrm2.tol, lrm3.95f),rounder), roc)
    
    #### The following code calcualte probabilities of occurrence
    binf <- cut(df2[,xvar], cutp, include.lowest = T)
    bvals <- tapply(df2[, isel] > 0, binf, mean, na.rm =T)
    
    if(plot.pdf) {
      ## Special Plot for log scale bc I can't get the original log scale to work properly 05212018
        if(emmaLOG==TRUE){
          file_name <-  paste('FinalDataAndPlots/Bug/',lim,'/',xvar,'/',tnames[i],'.jpg',sep='')
          jpeg(file_name)
          par(mar =mar, oma=c(1.5, 0.5, 2,0.5) )
          typ <- ifelse(samplen >= cutoff, "l", "n")
          if(xvar== "wshdImpPCT"){xrange[1] <- c(0.5)} # can't have 0 for log x scale ########################################
          plot(xnew, mean.resp, axes = TRUE, type = typ, xlab = "", ylab = "", xlim= xrange,
               ylim = range(low.bound, up.bound, bvals, na.rm =T), col=1, log='x')
          points(xnew, low.bound, type="l", col=1, lty="dashed")
          points(xnew, up.bound, type="l", col=1, lty="dashed")
          abline(v= lrm3.95f, lty= "dashed", col="red")
          y1range <- range(low.bound, up.bound, bvals, na.rm =T)
          y2range <- range(df1[, isel])
          if(add.abund) {
            y2new <- (df1[,isel] - min(y2range))*diff(y1range)/diff(y2range)+ min(y1range)    # convert y2 to y1 scale
            y1new.at <- (axTicks(2) - min(y1range))*diff(y2range)/diff(y1range)+ min(df1[, isel])    # convert to y2 scale
            y2.lab <- round(y1new.at, 3)    # add labels at original y2 scale
            axis(4, at = axTicks(2), labels = y2.lab, tcl =-0.4, lwd.ticks = 1)    # major ticks with labels
            points(dose, y2new, pch = 21, col = 1, cex = 0.3, bg = "lightgray")
            mtext(ylab2, side = 4, line = 2.5, col=1 , cex = 0.9)
          }    else {
            points(cutm, bvals, pch=21, col= 1, cex = 0.7, bg ="gray")                      # probability
          }
          mtext(xlabs, side = 1, line = 2.3, cex = 0.9)
          mtext(ylab1, side = 2, line = 2.3, cex = 0.9, col=1)
          mtext(bquote(italic(.(tnames[i]))), side = 3, line = 0.5)
          dev.off()
        }else{
          file_name <- paste('FinalDataAndPlots/Bug/',lim,'/',xvar,'/',tnames[i],'.jpg',sep='')
          jpeg(file_name)
          par(mar =mar, oma=c(1.5, 0.5, 2,0.5) )
          typ <- ifelse(samplen >= cutoff, "l", "n")
          plot(xnew, mean.resp, axes = FALSE, type = typ, xlab = "", ylab = "", xlim= xrange,
               ylim = range(low.bound, up.bound, bvals, na.rm =T), col=1)
          
          if(samplen >= cutoff) {
            points(xnew, low.bound, type="l", col=1, lty="dashed")
            points(xnew, up.bound, type="l", col=1, lty="dashed")
            if(!is.na(lim)) {  #  plot limits
              if(lim =="GAM") { 
                abline(v= lrm3.95f, lty= "dashed", col="red")
              } else if(lim =="CDF"){
                abline(v= eout2, lty= "dashed", col="blue")
              }
            }
          }   
          max.pow <- ceiling(max(xrange)); min.pow <- floor(min(xrange))  ## add x range for log formation    
          
          if(log.x) {
            at0 <- min.pow:max.pow              # add ticks at log10 = even     
            lab1 <- base^at0          # could be base = exp(1)
            axis(1, at = log(lab1 + plus, base), labels = lab1, lwd.ticks = 1)    # major ticks with labels
            xtick <- at0 <= max(df1[,xvar]) & at0 >= min(df1[,xvar])          # major labels
            
            if(sum(xtick) <5) {
              axis(1, at = log((1:10 * rep(lab1[-1]/10, each = 10)) + plus, base ), tcl = -0.4, 
                   labels = FALSE, lwd.ticks = 0.8)  # minor ticks            xtick <- at0 <= max(x) & axis.at >= xmin          # major labels
            }
            ### if only two or fewer major lables, then add tick labels at 2 and 5 log
            if(sum(xtick)==2) {
              at1 <- log(3 * rep(lab1[-1]/10, each = 1) + plus, base)
              axis(1, at = at1, tcl = -0.3, lwd.ticks= 1, labels = (3 * rep(lab1[-1]/10, each = 1)),
                   tcl = - 0.4)
            }   else if (sum(xtick)<2) {
              at2 <- log(c(2, 5) * rep(lab1[-1]/10, each = 2) + plus, base)
              axis(1, at = at2, tcl = -0.3, lwd.ticks= 1, labels = (c(2, 5) * 
                                                                      rep(lab1[-1]/10, each = 2)))
            }            
          } else  {
            axis(1, at = pretty(xrange), labels = pretty(xrange), lwd.ticks = 1, las = 1)    # major ticks with labels
          }
          axis(2)
          y1range <- range(low.bound, up.bound, bvals, na.rm =T)
          y2range <- range(df1[, isel])
          if(add.abund) {
            y2new <- (df1[,isel] - min(y2range))*diff(y1range)/diff(y2range)+ min(y1range)    # convert y2 to y1 scale
            y1new.at <- (axTicks(2) - min(y1range))*diff(y2range)/diff(y1range)+ min(df1[, isel])    # convert to y2 scale
            y2.lab <- round(y1new.at, 3)    # add labels at original y2 scale
            axis(4, at = axTicks(2), labels = y2.lab, tcl =-0.4, lwd.ticks = 1)    # major ticks with labels
            points(dose, y2new, pch = 21, col = 1, cex = 0.3, bg = "lightgray")
            mtext(ylab2, side = 4, line = 2.5, col=1 , cex = 0.9)
          }    else {
            points(cutm, bvals, pch=21, col= 1, cex = 0.7, bg ="gray")                      # probability
          }
          
          mtext(xlabs, side = 1, line = 2.3, cex = 0.9)
          mtext(ylab1, side = 2, line = 2.3, cex = 0.9, col=1)
          mtext(bquote(italic(.(tnames[i]))), side = 3, line = 0.5)
          box() 
          
          #if((i+5)/6-round((i+5)/6)==0) {
          #  title(main= main, outer=TRUE, cex.main= 1.8) 
          #  mtext(paste("Page", (i+5)/6, "of", totpage), outer = TRUE, side =1)
          #}
          
          dev.off()
        }
    }
      
    
    }
  }
  if(plot.pdf) graphics.off()
  return(optout <- data.frame(tnames, optsave))
}








tolerance_bioregion <- function(spdata, envdata,  sp.siteid="Sample.ID", species="GENUS", 
                      sp.abndid="RA", env.siteid="Sample.ID", xvar="cond", 
                      cutoff = 20, cutoff2 = 10, region = "all", lim ="GAM",
                      coord = NULL, mtype = 3, dense.N = 201, cast = TRUE,  
                      plot.pdf = F, add.map=F,statename = NULL,  add.lab = F, 
                      bioregion = 'Mountain',
                      add.abund = T, ylab1 = "Capture Probability",  
                      ylab2 = "Relative Abundance", 
                      main = "Capture Probability of Macroinvertebrate Taxon Along Conductivity Gradient", 
                      mar = c(5,4,3, 4), 
                      xlabs=expression(paste("Conductivity ( ", mu, "S/cm)")), 
                      log.x = TRUE, base = 10, plus = F, rounder = 0, taus = c(50,95), 
                      nbin = 60, emmaLOG = FALSE) {
  
  print(xvar)
  dfenv <- envdata[!is.na(envdata[,xvar]), c(env.siteid, xvar, coord)] 
  if(cast) { 
    df.sp <- data.frame(SITEID = spdata[, sp.siteid], species = spdata[, species],
                        abund = spdata[, sp.abndid]) 
    df.sp <- subset(df.sp, !is.na(species) & !is.na(abund))
    df.tmp <- merge(df.sp, dfenv, by.x = "SITEID", by.y = env.siteid )
    ss <- reshape::cast(df.tmp, SITEID ~ species, sum, value = "abund") 
    names(ss)[1] <- sp.siteid
  } else ss <- spdata
  #    print(names(ss))
  df1 <- merge(ss, dfenv, by.x = sp.siteid, by.y = env.siteid)   # merged crosstab abundance file
  nc <- ncol(df1)
  taxa.count <- apply(df1[,(2:ncol(ss))] > 0, 2, sum)
  #    print((taxa.count))
  tnames <- names(taxa.count)[taxa.count >= cutoff2] 
  ntaxa <- length(tnames)                           # length of taxa list for taxonomic level i
  
  df2 <- df1 <- df1[order(df1[,xvar]),]
  
  df2[2:(nc-ncol(dfenv) + 1)] <- apply(df2[2:(nc-ncol(dfenv) + 1 )] > 0, 2, as.numeric)                 # p/a file
  
  stress <- df1[, xvar]
  stress.u <- sort(unique(stress))
  
  xrange <- range(stress)
  xnew <- seq(from = xrange[1], to = xrange[2], length = 100)
  
  ###### calculate weight for cdf 
  
  cutp <- seq(from = min(stress), to = max(stress), length = nbin)
  cutm <- 0.5*(cutp[-1] + cutp[-nbin])
  
  df2$cutf <- cut(df2[,xvar], cutp, include.lowest = T)
  
  wt <- 1/table(df2$cutf)
  wt[is.infinite(wt)] <- NA
  wt <- wt/sum(wt, na.rm = T)
  dftemp <- data.frame(cutf = names(wt), wt = as.vector(wt))
  df2 <- merge(df2, dftemp, by = "cutf")[-1]
  df2 <- df2[order(df2[,xvar]),]  
  df1 <- df1[order(df1[,xvar]),] 
  
  #### end 
  ##### function to compute area under the curve
  auc <- function(xrange, mod, dense.N) {
    x <- seq(min(xrange), max(xrange), length = dense.N - 1)
    s.area <-rep(NA, dense.N - 1)
    y <- predict(mod, newdata = data.frame(dose = x), type = "response")
    for (index in 1:dense.N-1) {
      s.area[index] <- (y[index] + y[index + 1])/2*(x[index+1]-x[index])
    }
    tsum <- sum(s.area, na.rm=T)         # all area
    jj=1
    csum = sum(s.area[1:jj])              # cum area
    xc95 <- yc95 <- rep(NA, length(taus))
    for(ii in 1:length(taus)) {
      while(csum < taus[ii]/100*tsum) {
        jj = jj + 1
        csum <- sum(s.area[1:jj], na.rm=T)
      }
      if(jj == 1) {
        xc95[ii] <- x[jj]
        yc95[ii] <- y[jj]
      } else {
        xc95[ii] <- (x[jj]+ x[jj-1])/2  
        yc95[ii] <- (y[jj]+ y[jj-1])/2
      }
    }
    return(c(xc95))
    #      print(xc95)
  } 
  # plot pdf option
  #if(plot.pdf) {
  #  pdf(file = paste("data/", region, ".taxon.gam.pdf",sep=""),
  #      width = 9, height = 6.5, pointsize = 12)
  #  par(mfrow = c(2, 3), pty = "m", mar =mar, oma=c(1.5, 0.5, 2,0.5) )
  #  totpage <- ceiling((length(tnames)+1)/6)
  #  
  #  if(add.map) {
  #    simpleCap <- function(x) {
  #      s <- strsplit(x, " ")[[1]]
  #      paste(toupper(substring(s, 1,1)), substring(s, 2),
  #            sep="", collapse=" ")
  #    }
  #    med = 1
  #    while(is.na(df1[med, coord[1]])) med = 1 + med
  #    
  #    if(is.null(statename)) {
  #      statename <<- map.where("state", df1[med, coord[1]], df1[med, coord[2]])
  #      statename <- simpleCap(statename)
  #    } 
  #    if(add.lab) {
  #      map.text("state", region = statename, mar = mar)
  #    } else {
  #      map("state", regions = statename, mar = mar)
  #    }
  #    
  #    text(df1[,coord[1]],df1[,coord[2]],".")
  #    #       mtext(paste("Ecoregion",region), side =3,line = 0.5, cex=1.5)  
  #  }
  #}
  
  varnames <- c("N", "min_ob", paste(taus, "th_ob",sep="_"), "max_ob", "Opt_WA",
                "Tol_WA", paste("CDF", taus,"th","Abund", sep ="_"),
                paste("CDF", taus,"th","PA", sep ="_"),paste("CDF_wt", taus,"th",sep ="_"), 
                paste("LRM", taus,"th", sep ="_"),paste("QLRM", taus,"th", sep ="_"),
                "Opt_qlrm", "Tol_qlrm", paste("GAM",taus, "th", sep="_"), "ROC")
  
  optsave <- matrix(NA, ncol = length(taus)*7 + 8, nrow  = ntaxa)
  colnames(optsave) <- varnames
  rownames(optsave) <- tnames
  
  for (i in 1:ntaxa) {
    isel <- match(tnames[i], names(df1))       # selected taxa i
    #    print(df1[,isel])
    print(tnames[i])
    ### (0) observed range
    samplen <- length(df2[df2[,isel]>0,xvar])          
    limits <- quantile(df2[df2[,isel]>0,xvar], prob = c(0, taus/100, 1))       
    
    ## (1)  Weighted averaging 
    WA <- sum(df1[,isel]*df1[,xvar])/sum(df1[,isel]) 
    tol <- sqrt(sum(df1[,isel] * (df1[,xvar]- WA)^2) /sum(df1[,isel])) 
    #    WTA <- sum(df1[,isel]*df1[,xvar]*df2$wt)/sum(df1[,isel]*df2$wt)     # sample weighted weighted average         
    
    ### (2)  cdf or cumsum no weighting  both abundance based and p/a based
    csum1 <-cumsum(df1[,isel])/sum(df1[,isel])          # cumlative vectors devided total abundance 
    #      print(sum(df1[,isel]))
    ic1 <- rep(1, length(taus))
    for(ii in 1:length(taus)) {
      #  print(csum1)
      if(is.finite(csum1[ic1[ii]])) {
        while(csum1[ic1[ii]] < taus[ii]/100) ic1[ii] <- ic1[ii] + 1
      } else csum1 <- NA
    }
    #### (3) cdf p/a        
    csum2 <-cumsum(df2[,isel])/sum(df2[,isel])    # cumlative vectors devided total abundance 
    ic2 <- rep(1, length(taus))
    for(ii in 1:length(taus)) {
      while(csum1[ic2[ii]] < taus[ii]/100) ic2[ii]<- ic2[ii] + 1
    }
    
    ##### (4) cdf weighted 
    pres <- df2[, isel] > 0
    sel <- df2[pres,]
    eout2 <- rep(NA, length(taus))
    for(ii in 1:length(taus)) {
      eout2[ii] <- Hmisc::wtd.quantile(sel[,xvar], sel$wt, normwt = TRUE, prob = taus[ii]/100)  
    }
    
    ######### (5)(6)(7) logistic regression model          
    resp <<- df2[, isel] 
    dose <<- df2[, xvar]
    lrm1 <- glm(resp ~ dose, family="binomial")
    lrm2 <- glm(resp ~ dose + I(dose^2), family="binomial")
    #### u = -b1/(2b2)  t = 1/sqrt(-2b2) for lrm2  
    lrm2.opt <- (-lrm2$coef[2]/(2*lrm2$coef[3])  )
    lrm2.tol <- 1/sqrt(-2*as.vector(lrm2$coef[3]))
    s=mgcv:::s
    
    lrm3 <- try(mgcv::gam(resp ~ s(dose,k = 3), family= "binomial"),TRUE)
    
    if(isTRUE(class(lrm3)=="try-error")) { next } else{  ############################ ERROR HANDLING INSERTED BY EVJ 02062018
    
    if(mtype == 1) model <- lrm1
    if(mtype == 2) model <- lrm2
    if(mtype == 3) model <- lrm3

    # Compute mean predicted probability of occurrence
    predlk <- predict(model, type = "link")
    predout <- exp(predlk)/(1 + exp(predlk))
    x <- predout[pres]
    y <- predout[!pres]
    rocmat <- matrix(NA, nrow = length(x), ncol = length(y))
    for (j in 1:length(x)) {
      rocmat[j,] <- as.numeric(x[j] > y)
    }
    # Summarize all comparisons to compute area under ROC
    roc <- sum(rocmat)/(length(x)*length(y))
    #        wilcox <- wilcox.test(x, y, conf.int = FALSE)
    predresp <- predict(model, newdata = data.frame(dose=xnew), type="link", se.fit=T)
    
    # Compute upper and lower 90% confidence limits
    up.bound.link <- predresp$fit + qnorm(0.95)*predresp$se.fit
    low.bound.link <- predresp$fit + qnorm(0.05)*predresp$se.fit
    mean.resp.link <- predresp$fit
    
    # Convert from logit transformed values to probability.
    up.bound <- exp(up.bound.link)/(1+exp(up.bound.link))
    low.bound <- exp(low.bound.link)/(1+exp(low.bound.link))
    mean.resp <- exp(mean.resp.link)/(1+exp(mean.resp.link))
    
    #### 5,6,7 full range
    ######## find modeled xc95 full data range
    lrm1.95f <- auc(xrange = xrange, mod = lrm1, dense.N = dense.N)
    lrm2.95f <- auc(xrange = xrange, mod = lrm2, dense.N = dense.N)
    lrm3.95f <- auc(xrange = xrange, mod = lrm3, dense.N = dense.N) 
    #        print(lrm3.95f)
    ###### Save results 
    
    optsave[i,] <- c(samplen, round(c(limits, WA, tol, df1[ic1,xvar],
                                      df1[ic2,xvar], eout2, lrm1.95f,lrm2.95f, lrm2.opt, lrm2.tol, lrm3.95f),rounder), roc)
    
    #### The following code calcualte probabilities of occurrence
    binf <- cut(df2[,xvar], cutp, include.lowest = T)
    bvals <- tapply(df2[, isel] > 0, binf, mean, na.rm =T)
    
    if(plot.pdf) {
      ## Special Plot for log scale bc I can't get the original log scale to work properly 05212018
        if(emmaLOG==TRUE){
          file_name <- paste('FinalDataAndPlots/Bug/Bioregion/',bioregion,'/',lim,'/',xvar,'/',tnames[i],'.jpg',sep='')
          jpeg(file_name)
          par(mar =mar, oma=c(1.5, 0.5, 2,0.5) )
          typ <- ifelse(samplen >= cutoff, "l", "n")
          if(xvar== "wshdImpPCT"){xrange[1] <- c(0.5)} # can't have 0 for log x scale
          plot(xnew, mean.resp, axes = TRUE, type = typ, xlab = "", ylab = "", xlim= xrange,
               ylim = range(low.bound, up.bound, bvals, na.rm =T), col=1, log='x')
          points(xnew, low.bound, type="l", col=1, lty="dashed")
          points(xnew, up.bound, type="l", col=1, lty="dashed")
          abline(v= lrm3.95f, lty= "dashed", col="red")
          y1range <- range(low.bound, up.bound, bvals, na.rm =T)
          y2range <- range(df1[, isel])
          if(add.abund) {
            y2new <- (df1[,isel] - min(y2range))*diff(y1range)/diff(y2range)+ min(y1range)    # convert y2 to y1 scale
            y1new.at <- (axTicks(2) - min(y1range))*diff(y2range)/diff(y1range)+ min(df1[, isel])    # convert to y2 scale
            y2.lab <- round(y1new.at, 3)    # add labels at original y2 scale
            axis(4, at = axTicks(2), labels = y2.lab, tcl =-0.4, lwd.ticks = 1)    # major ticks with labels
            points(dose, y2new, pch = 21, col = 1, cex = 0.3, bg = "lightgray")
            mtext(ylab2, side = 4, line = 2.5, col=1 , cex = 0.9)
          }    else {
            points(cutm, bvals, pch=21, col= 1, cex = 0.7, bg ="gray")                      # probability
          }
          mtext(xlabs, side = 1, line = 2.3, cex = 0.9)
          mtext(ylab1, side = 2, line = 2.3, cex = 0.9, col=1)
          mtext(bquote(italic(.(tnames[i]))), side = 3, line = 0.5)
          dev.off()
        }else{
          file_name <- paste('FinalDataAndPlots/Bug/Bioregion/',bioregion,'/',lim,'/',xvar,'/',tnames[i],'.jpg',sep='')
          jpeg(file_name)
          par(mar =mar, oma=c(1.5, 0.5, 2,0.5) )
          typ <- ifelse(samplen >= cutoff, "l", "n")
          plot(xnew, mean.resp, axes = FALSE, type = typ, xlab = "", ylab = "", xlim= xrange,
               ylim = range(low.bound, up.bound, bvals, na.rm =T), col=1)
          
          if(samplen >= cutoff) {
            points(xnew, low.bound, type="l", col=1, lty="dashed")
            points(xnew, up.bound, type="l", col=1, lty="dashed")
            if(!is.na(lim)) {  #  plot limits
              if(lim =="GAM") { 
                abline(v= lrm3.95f, lty= "dashed", col="red")
              } else if(lim =="CDF"){
                abline(v= eout2, lty= "dashed", col="blue")
              }
            }
          }   
          max.pow <- ceiling(max(xrange)); min.pow <- floor(min(xrange))  ## add x range for log formation    
          
          if(log.x) {
            at0 <- min.pow:max.pow              # add ticks at log10 = even     
            lab1 <- base^at0          # could be base = exp(1)
            axis(1, at = log(lab1 + plus, base), labels = lab1, lwd.ticks = 1)    # major ticks with labels
            xtick <- at0 <= max(df1[,xvar]) & at0 >= min(df1[,xvar])          # major labels
            
            if(sum(xtick) <5) {
              axis(1, at = log((1:10 * rep(lab1[-1]/10, each = 10)) + plus, base ), tcl = -0.4, 
                   labels = FALSE, lwd.ticks = 0.8)  # minor ticks            xtick <- at0 <= max(x) & axis.at >= xmin          # major labels
            }
            ### if only two or fewer major lables, then add tick labels at 2 and 5 log
            if(sum(xtick)==2) {
              at1 <- log(3 * rep(lab1[-1]/10, each = 1) + plus, base)
              axis(1, at = at1, tcl = -0.3, lwd.ticks= 1, labels = (3 * rep(lab1[-1]/10, each = 1)),
                   tcl = - 0.4)
            }   else if (sum(xtick)<2) {
              at2 <- log(c(2, 5) * rep(lab1[-1]/10, each = 2) + plus, base)
              axis(1, at = at2, tcl = -0.3, lwd.ticks= 1, labels = (c(2, 5) * 
                                                                      rep(lab1[-1]/10, each = 2)))
            }            
          } else  {
            axis(1, at = pretty(xrange), labels = pretty(xrange), lwd.ticks = 1, las = 1)    # major ticks with labels
          }
          axis(2)
          y1range <- range(low.bound, up.bound, bvals, na.rm =T)
          y2range <- range(df1[, isel])
          if(add.abund) {
            y2new <- (df1[,isel] - min(y2range))*diff(y1range)/diff(y2range)+ min(y1range)    # convert y2 to y1 scale
            y1new.at <- (axTicks(2) - min(y1range))*diff(y2range)/diff(y1range)+ min(df1[, isel])    # convert to y2 scale
            y2.lab <- round(y1new.at, 3)    # add labels at original y2 scale
            axis(4, at = axTicks(2), labels = y2.lab, tcl =-0.4, lwd.ticks = 1)    # major ticks with labels
            points(dose, y2new, pch = 21, col = 1, cex = 0.3, bg = "lightgray")
            mtext(ylab2, side = 4, line = 2.5, col=1 , cex = 0.9)
          }    else {
            points(cutm, bvals, pch=21, col= 1, cex = 0.7, bg ="gray")                      # probability
          }
          
          mtext(xlabs, side = 1, line = 2.3, cex = 0.9)
          mtext(ylab1, side = 2, line = 2.3, cex = 0.9, col=1)
          mtext(bquote(italic(.(tnames[i]))), side = 3, line = 0.5)
          box() 
          
          #if((i+5)/6-round((i+5)/6)==0) {
          #  title(main= main, outer=TRUE, cex.main= 1.8) 
          #  mtext(paste("Page", (i+5)/6, "of", totpage), outer = TRUE, side =1)
          #}
          
          dev.off()
        }
    }
    }
  }
  if(plot.pdf) graphics.off()
  return(optout <- data.frame(tnames, optsave))
}
```


And then one more function to help make with plot automation.
```{r, echo=F}

# function to get better x axis label
prettyXlabel <- function(parameter){
  if(parameter=="DO"){return('Dissolved Oxygen (mg/L)')}
  if(parameter=="pH"){return('pH (Standard Units)')}
  if(parameter=="SpCond"){return(expression(paste("Specific Conductivity ( ", mu, "S/cm)")))}
  if(parameter=="TN"){return('Total Nitrogen (mg/L)')}
  if(parameter=="TP"){return('Total Phosphorus (mg/L)')}
  if(parameter=='TDS'){return('Total Dissolved Solids (mg/L)')}
  if(parameter=="NH4"){return('Ammonia Nitrogen (mg/L)')}
  if(parameter=="NO3"){return('Total Nitrate Nitrogen (mg/L)')}
  if(parameter=="TKN"){return('Total Kjendahl Nitrogen (mg/L)')}
  if(parameter=="Ortho-P"){return("Total Ortho-Phosphorus (mg/L)")}
  if(parameter== "Turb" ){return('Turbidity (NTU)')}
  if(parameter== "TSS" ){return('Total Suspended Solids (mg/L)')}
  if(parameter=="Na"){return("Dissolved Sodium (mg/L)")}
  if(parameter=="K"){return("Dissolved Potassium (mg/L)")}
  if(parameter=="Cl"){return("Dissolved Chloride (mg/L)")}
  if(parameter=="Sf"){return("Dissolved Sulfate (mg/L)")}
  if(parameter=="70331VFine"){return("% Fine Sediment")}
  if(parameter=="SSCCOARSE"){return("%SSC Coarse")}
  if(parameter=="SSCFINE"){return("%SSC Fine")}
  if(parameter=="SSCTOTAL" ){return("SSC Course + SSC Fine Combined")}
  if(parameter=="LRBS"){return("Logged Relative Bed Stability")}
  if(parameter=="Slope"){return("% Slope")} 
  if(parameter=="FN_PCT"){return("% Fine Sediment")}
  if(parameter=="SA_PCT"){return("% Sand")}
  if(parameter=="SA_FN_PCT"){return("%Fine + %Sand")}
  if(parameter=="LSUB_DMM"){return("Logged Mean Particle Size in Reach")}
  if(parameter=="BL_CB_GR_Embed_PCT"){return("% Embeddedness of Bolders,Cobbles, and Gravels")}
  if(parameter=="Embed_PCT"){return("% Embeddedness All Particles")}
  if(parameter=="TotHab"){return("Total Habitat Score (RBP) ")}
  if(parameter=="TotHab_wMD"){return("Total Habitat Score (with MD) (RBP_MD conversion) ")}
  if(parameter=="MetalCCU"){return("Metal chronic criterion unit (Index of multiple metals)")}
}

```



### Storing the data and plots
Now that the functions are built to roll through making CDF and GAM plots for each taxa across various stressors, we can run through stressors to create plots and bring into Attribution workshop shiny app. The CDF data will be run but the plots will be generated using plotly in the app for interactive features. TITAN plots will be stored as flat files like the GAM/rel abundance.
*Note* One must make necessary file structure in Windows explorer first for the functions to behave as expected.
```{r}

# Save CDF data
cdfdataDO <- CDF(merge(bugTaxa,select(envData,UID,DO),by='UID') %>%
                   filter(!is.na(DO)),'DO')
cdfdataDO_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,DO),by='UID') %>%
                           filter(!is.na(DO)),'DO')
cdfdataDO_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,DO),by='UID') %>%
                           filter(!is.na(DO)),'DO')
                                  
cdfdataDO_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,DO),by='UID') %>%
                           filter(!is.na(DO)),'DO')
saveRDS(cdfdataDO,'FinalDataAndPlots/Bug/CDF/cdfdataDO.RDS')
saveRDS(cdfdataDO_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataDO.RDS')
saveRDS(cdfdataDO_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataDO.RDS')
saveRDS(cdfdataDO_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataDO.RDS')

cdfdatapH <- CDF(merge(bugTaxa,select(envData,UID,pH),by='UID') %>%
                   filter(!is.na(pH)),'pH')
cdfdatapH_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,pH),by='UID') %>%
                           filter(!is.na(pH)),'pH')
cdfdatapH_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,pH),by='UID') %>%
                           filter(!is.na(pH)),'pH')
                                  
cdfdatapH_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,pH),by='UID') %>%
                           filter(!is.na(pH)),'pH')
saveRDS(cdfdatapH,'FinalDataAndPlots/Bug/CDF/cdfdatapH.RDS')
saveRDS(cdfdatapH_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdatapH.RDS')
saveRDS(cdfdatapH_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdatapH.RDS')
saveRDS(cdfdatapH_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdatapH.RDS')

cdfdataSpCond <- CDF(merge(bugTaxa,select(envData,UID,SpCond),by='UID') %>%
                   filter(!is.na(SpCond)),'SpCond')
cdfdataSpCond_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,SpCond),by='UID') %>%
                           filter(!is.na(SpCond)),'SpCond')
cdfdataSpCond_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,SpCond),by='UID') %>%
                           filter(!is.na(SpCond)),'SpCond')
                                  
cdfdataSpCond_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,SpCond),by='UID') %>%
                           filter(!is.na(SpCond)),'SpCond')
saveRDS(cdfdataSpCond,'FinalDataAndPlots/Bug/CDF/cdfdataSpCond.RDS')
saveRDS(cdfdataSpCond_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataSpCond.RDS')
saveRDS(cdfdataSpCond_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataSpCond.RDS')
saveRDS(cdfdataSpCond_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataSpCond.RDS')
                         
cdfdataLRBS <- CDF(merge(bugTaxa,select(envData,UID,LRBS),by='UID') %>%
                   filter(!is.na(LRBS)),'LRBS')
cdfdataLRBS_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,LRBS),by='UID') %>%
                           filter(!is.na(LRBS)),'LRBS')
cdfdataLRBS_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,LRBS),by='UID') %>%
                           filter(!is.na(LRBS)),'LRBS')
                                  
cdfdataLRBS_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,LRBS),by='UID') %>%
                           filter(!is.na(LRBS)),'LRBS')
saveRDS(cdfdataLRBS,'FinalDataAndPlots/Bug/CDF/cdfdataLRBS.RDS')
saveRDS(cdfdataLRBS_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataLRBS.RDS')
saveRDS(cdfdataLRBS_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataLRBS.RDS')
saveRDS(cdfdataLRBS_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataLRBS.RDS')


cdfdataTP <- CDF(merge(bugTaxa,select(envData,UID,TP),by='UID') %>%
                   filter(!is.na(TP)),'TP')
cdfdataTP_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,TP),by='UID') %>%
                           filter(!is.na(TP)),'TP')
cdfdataTP_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,TP),by='UID') %>%
                           filter(!is.na(TP)),'TP')
                                  
cdfdataTP_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,TP),by='UID') %>%
                           filter(!is.na(TP)),'TP')
saveRDS(cdfdataTP,'FinalDataAndPlots/Bug/CDF/cdfdataTP.RDS')
saveRDS(cdfdataTP_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataTP.RDS')
saveRDS(cdfdataTP_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataTP.RDS')
saveRDS(cdfdataTP_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataTP.RDS')

cdfdataTN <- CDF(merge(bugTaxa,select(envData,UID,TN),by='UID') %>%
                   filter(!is.na(TN)),'TN')
cdfdataTN_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,TN),by='UID') %>%
                           filter(!is.na(TN)),'TN')
cdfdataTN_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,TN),by='UID') %>%
                           filter(!is.na(TN)),'TN')
                                  
cdfdataTN_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,TN),by='UID') %>%
                           filter(!is.na(TN)),'TN')
saveRDS(cdfdataTN,'FinalDataAndPlots/Bug/CDF/cdfdataTN.RDS')
saveRDS(cdfdataTN_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataTN.RDS')
saveRDS(cdfdataTN_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataTN.RDS')
saveRDS(cdfdataTN_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataTN.RDS')

cdfdataTotHab <- CDF(merge(bugTaxa,select(envData,UID,TotHab),by='UID') %>%
                   filter(!is.na(TotHab)),'TotHab')
cdfdataTotHab_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,TotHab),by='UID') %>%
                           filter(!is.na(TotHab)),'TotHab')
cdfdataTotHab_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,TotHab),by='UID') %>%
                           filter(!is.na(TotHab)),'TotHab')
                                  
cdfdataTotHab_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,TotHab),by='UID') %>%
                           filter(!is.na(TotHab)),'TotHab')
saveRDS(cdfdataTotHab,'FinalDataAndPlots/Bug/CDF/cdfdataTotHab.RDS')
saveRDS(cdfdataTotHab_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataTotHab.RDS')
saveRDS(cdfdataTotHab_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataTotHab.RDS')
saveRDS(cdfdataTotHab_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataTotHab.RDS')

cdfdatawshdImpPCT <- CDF(merge(bugTaxa,select(envData,UID,wshdImpPCT),by='UID') %>%
                   filter(!is.na(wshdImpPCT)),'wshdImpPCT')
cdfdatawshdImpPCT_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,wshdImpPCT),by='UID') %>%
                           filter(!is.na(wshdImpPCT)),'wshdImpPCT')
cdfdatawshdImpPCT_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,wshdImpPCT),by='UID') %>%
                           filter(!is.na(wshdImpPCT)),'wshdImpPCT')
                                  
cdfdatawshdImpPCT_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,wshdImpPCT),by='UID') %>%
                           filter(!is.na(wshdImpPCT)),'wshdImpPCT')
saveRDS(cdfdatawshdImpPCT,'FinalDataAndPlots/Bug/CDF/cdfdatawshdImpPCT.RDS')
saveRDS(cdfdatawshdImpPCT_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdatawshdImpPCT.RDS')
saveRDS(cdfdatawshdImpPCT_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdatawshdImpPCT.RDS')
saveRDS(cdfdatawshdImpPCT_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdatawshdImpPCT.RDS')

cdfdataTotHab_wMD <- CDF(merge(bugTaxa,select(envData,UID,TotHab_wMD),by='UID') %>%
                   filter(!is.na(TotHab_wMD)),'TotHab_wMD')
cdfdataTotHab_wMD_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,TotHab_wMD),by='UID') %>%
                           filter(!is.na(TotHab_wMD)),'TotHab_wMD')
cdfdataTotHab_wMD_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,TotHab_wMD),by='UID') %>%
                           filter(!is.na(TotHab_wMD)),'TotHab_wMD')
                                  
cdfdataTotHab_wMD_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,TotHab_wMD),by='UID') %>%
                           filter(!is.na(TotHab_wMD)),'TotHab_wMD')
saveRDS(cdfdataTotHab_wMD,'FinalDataAndPlots/Bug/CDF/cdfdataTotHab_wMD.RDS')
saveRDS(cdfdataTotHab_wMD_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataTotHab_wMD.RDS')
saveRDS(cdfdataTotHab_wMD_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataTotHab_wMD.RDS')
saveRDS(cdfdataTotHab_wMD_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataTotHab_wMD.RDS')

cdfdataSf <- CDF(merge(bugTaxa,select(envData,UID,Sf),by='UID') %>%
                   filter(!is.na(Sf)),'Sf')
cdfdataSf_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,Sf),by='UID') %>%
                           filter(!is.na(Sf)),'Sf')
cdfdataSf_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,Sf),by='UID') %>%
                           filter(!is.na(Sf)),'Sf')
                                  
cdfdataSf_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,Sf),by='UID') %>%
                           filter(!is.na(Sf)),'Sf')
saveRDS(cdfdataSf,'FinalDataAndPlots/Bug/CDF/cdfdataSf.RDS')
saveRDS(cdfdataSf_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataSf.RDS')
saveRDS(cdfdataSf_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataSf.RDS')
saveRDS(cdfdataSf_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataSf.RDS')

cdfdataCl <- CDF(merge(bugTaxa,select(envData,UID,Cl),by='UID') %>%
                   filter(!is.na(Cl)),'Cl')
cdfdataCl_Mountain <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Mountain') %>%
                                  select(UID,Cl),by='UID') %>%
                           filter(!is.na(Cl)),'Cl')
cdfdataCl_Piedmont <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Piedmont') %>%
                                  select(UID,Cl),by='UID') %>%
                           filter(!is.na(Cl)),'Cl')
                                  
cdfdataCl_Coast <- CDF(merge(bugTaxa,filter(envData,BioRegion=='Coast') %>%
                               select(UID,Cl),by='UID') %>%
                           filter(!is.na(Cl)),'Cl')
saveRDS(cdfdataCl,'FinalDataAndPlots/Bug/CDF/cdfdataCl.RDS')
saveRDS(cdfdataCl_Mountain,'FinalDataAndPlots/Bug/Bioregion/Mountain/CDF/cdfdataCl.RDS')
saveRDS(cdfdataCl_Piedmont,'FinalDataAndPlots/Bug/Bioregion/Piedmont/CDF/cdfdataCl.RDS')
saveRDS(cdfdataCl_Coast,'FinalDataAndPlots/Bug/Bioregion/Coast/CDF/cdfdataCl.RDS')






parameters <- c("DO","pH","SpCond","LRBS","TP","TN","wshdImpPCT","TotHab",'TotHab_wMD','Sf','Cl')#,"MetalCCU")
parametersToAnalyze <- which( colnames(envData) %in% parameters)

# GAM/relative Risk plots
for(i in parametersToAnalyze){
  z <- tolerance(spdata=gather(bugTaxa,GENUS,RA,-UID),
                 envdata = select(envData,UID,names(envData)[i]),
                 sp.siteid="UID" ,species="GENUS",sp.abndid="RA",env.siteid="UID",
                 xvar=as.character(names(envData)[i]),
                 cutoff = 10,
                 cutoff2 = 10,
                 region = "all",lim ="GAM",coord = NULL,mtype = 3,dense.N = 201,
                 cast = TRUE  ,plot.pdf = T,add.map=F,statename = NULL,
                 add.lab = F,add.abund = T,
                 ylab1 = "Capture Probability",ylab2 = "Relative Abundance",
                 main = "Capture Probability of Bug Taxon Along Conductivity Gradient",
                 mar = c(5,4,3, 4),
                 xlabs=prettyXlabel(names(envData)[i]),
                 log.x = FALSE,
                 base = 10,plus = F,rounder = 2,taus = c(50,95),nbin = 60,
                 emmaLOG = ifelse(i %in% c(14,15,16,19,20,23),TRUE,FALSE))
  saveRDS(z,paste('FinalDataAndPlots/Bug/GAM/',as.character(names(envData)[i]),'/',
                  as.character(names(envData)[i]),'.RDS',sep=''))
  
  # Mountain
  y <- tolerance_bioregion(spdata=gather(bugTaxa,GENUS,RA,-UID),
                           envdata = filter(envData, BioRegion=='Mountain') %>%
                             select(UID,names(envData)[i]) %>% filter(!is.na(.[[2]])),
                 sp.siteid="UID" ,species="GENUS",sp.abndid="RA",env.siteid="UID",
                 xvar=as.character(names(envData)[i]),
                 cutoff = 10,
                 cutoff2 = 10,
                 region = "all",lim ="GAM",coord = NULL,mtype = 3,dense.N = 201,
                 cast = TRUE  ,plot.pdf = T,add.map=F,statename = NULL,
                 add.lab = F,
                 
                 bioregion = 'Mountain',
                 
                 add.abund = T,
                 ylab1 = "Capture Probability",ylab2 = "Relative Abundance",
                 main = "Capture Probability of Bug Taxon Along Conductivity Gradient",
                 mar = c(5,4,3, 4),
                 xlabs=prettyXlabel(names(envData)[i]),
                 log.x = FALSE,
                 base = 10,plus = F,rounder = 2,taus = c(50,95),nbin = 60,
                 emmaLOG = ifelse(i %in% c(14,15,16,19,20,23),TRUE,FALSE))
                 
  saveRDS(y,paste('FinalDataAndPlots/Bug/Bioregion/Mountain/GAM/',as.character(names(envData)[i]),'/',
                  as.character(names(envData)[i]),'.RDS',sep=''))
  
  
   # Piedmont
  x <- tolerance_bioregion(spdata=gather(bugTaxa,GENUS,RA,-UID),
                           envdata = filter(envData, BioRegion=='Piedmont') %>%
                             select(UID,names(envData)[i]) %>% filter(!is.na(.[[2]])),
                 sp.siteid="UID" ,species="GENUS",sp.abndid="RA",env.siteid="UID",
                 xvar=as.character(names(envData)[i]),
                 cutoff = 10,
                 cutoff2 = 10,
                 region = "all",lim ="GAM",coord = NULL,mtype = 3,dense.N = 201,
                 cast = TRUE  ,plot.pdf = T,add.map=F,statename = NULL,
                 add.lab = F,
                 
                 bioregion = 'Piedmont',
                 
                 add.abund = T,
                 ylab1 = "Capture Probability",ylab2 = "Relative Abundance",
                 main = "Capture Probability of Bug Taxon Along Conductivity Gradient",
                 mar = c(5,4,3, 4),
                 xlabs=prettyXlabel(names(envData)[i]),
                 log.x = FALSE,
                 base = 10,plus = F,rounder = 2,taus = c(50,95),nbin = 60,
                 emmaLOG = ifelse(i %in% c(14,15,16,19,20,23),TRUE,FALSE))
                 
  saveRDS(x,paste('FinalDataAndPlots/Bug/Bioregion/Piedmont/GAM/',as.character(names(envData)[i]),'/',
                  as.character(names(envData)[i]),'.RDS',sep=''))
  
  # Coast
  w <- tolerance_bioregion(spdata=gather(bugTaxa,GENUS,RA,-UID),
                           envdata = filter(envData, BioRegion=='Coast') %>%
                             select(UID,names(envData)[i]) %>% filter(!is.na(.[[2]])),
                 sp.siteid="UID" ,species="GENUS",sp.abndid="RA",env.siteid="UID",
                 xvar=as.character(names(envData)[i]),
                 cutoff = 10,
                 cutoff2 = 10,
                 region = "all",lim ="GAM",coord = NULL,mtype = 3,dense.N = 201,
                 cast = TRUE  ,plot.pdf = T,add.map=F,statename = NULL,
                 add.lab = F,
                 
                 bioregion = 'Coast',
                 
                 add.abund = T,
                 ylab1 = "Capture Probability",ylab2 = "Relative Abundance",
                 main = "Capture Probability of Bug Taxon Along Conductivity Gradient",
                 mar = c(5,4,3, 4),
                 xlabs=prettyXlabel(names(envData)[i]),
                 log.x = FALSE,
                 base = 10,plus = F,rounder = 2,taus = c(50,95),nbin = 60,
                 emmaLOG = ifelse(i %in% c(14,15,16,19,20,23),TRUE,FALSE))
                 
  saveRDS(w,paste('FinalDataAndPlots/Bug/Bioregion/Coast/GAM/',as.character(names(envData)[i]),'/',
                  as.character(names(envData)[i]),'.RDS',sep=''))
}

```